Scheduling and Amortization for MPC

### 2022-05-25 notes/notes.tex
Commit before clean up.

>>>>

\section{Array SSA Algorithm}

\begin{verbatim}
// initialize the downwardExposed map for entryBlock with the original A's.
// ANA: I think this will be the phi-node block when extending with nested loops
add entryBlock to WL
BFS(WL)

BFS(worklist) // bfs traversal of loop body CFG
  block <- remove next basic block from WL queue
  process(block) 
  for each block1 in succ(block), add block1 to WL if not already there.

process(basicBlock) 
  if basicBlock is a merge block 
    B1 <- predecessor of basicBlock
    B2 <- predecessor of basicBlock
    for each array A
       currentA_B1 <- downwardExposed.get(B1).get(A) 
       currentA_B2 <- downwardExposed.get(B2).get(A) 
       if currentA_B1 != currentA_B2 
       // if there were modifications in either B1 or B2 will differ
          newA <- phi(currentAB1, currentAB2) // creates phi node for array A
          downwardExposed.get(basicBlock).put(A,newA) 
  else // it is not a merge block
    Block <- predecessor of basicBlock // single pred, not a merge node
    for each array A
       currentA <- downwardExposed.get(Block).get(A)
       downwardExposed.get(basicBlock).put(A,currentA)

  // now start the traversal of statements in basicBlock:
  for (each stmt in statements of basicBlock)
     case stmt is array write A[f(i)] = ...
         newA <- create a new version of A
         downwardExposed.get(basicBlock).put(A,newA) 
         // have to add the new stmt
     case stmt is array read ... = A[f(i)]
         currentA <- downwardExposed.get(basicBlock).get(A)
         change stmt to ... = currentA[f(i)]

\end{verbatim}

### 2022-05-25 paper.tex
Commit before clean up.

@@

\begin{figure}
\small
\[
\begin{array}{l@{~~~~~}l}
  \begin{array}{l@{~}l@{~~~}l}
  s & ::= s ; s \\ 
  & \mid \x = \y \; \mid \x=\y \; \op \; \z \mid {\sf A[f(i)] = x} \; \mid {\sf x = A[f(i)]} & \mathit{assignment} \\
%  & \mid \x = \phi(\y,\z) \\
  & \mid {\sf for} \; (i = 0; i \le \mathit{Const}; \mathit{i\!+\!+}) \; \{ \; s \; \} \\
  & \mid {\sf if} \; (\x \; \cop \; \y) \; \{ \; s \; \} \; {\sf else} \; \{ \; s \; \} \\
 % \bool &::= \x \; \cop \; \y & \mathit{comparison} \\
  \op &::= + \mid - \mid * \mid / & \mathit{arithmetic} \; \mathit{operator} \\
  \cop &::=  \; == \; \mid \; != \; \mid \; < \; \mid \; \le & \mathit{comparison} \; \mathit{operator}

@@

\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
z = 0;
...
if (x > 0) {
  if (y > 0) {
    z = 1;
  }  else {
    z = -1;
  }
}
print z
\end{minted}
Line 10 is the dominance frontier of the assignments to \z, and Cytron et. al.'s algorithm
creates a $\phi$ node at 10: $z_3 = \phi(z_0, z_1, z_2)$. There is no information retained
that $z_0$ is triggered when $x \le 0$, $z_1$ is triggered when $x > 0 \; \wedge \; y > 0$, 
and $z_2$ is triggered when $x > 0 \; \wedge \; y \le 0$. The Multiplexer at the join point
needs this information however in order to assign the correct value to $z_3$ at runtime.

@@

\begin{algorithm}
\caption{Determine Graph edges}
\begin{algorithmic}[1]
%\REQUIRE $c \neq null$
%\STATE $L \leftarrow get\_nodes(c)$
%\FORALL{$n$ in $L$}
%\STATE $...$
%\ENDFOR
\FORALL {variables $V$}
\STATE $C(V) \leftarrow 0$
\STATE $S(V) \leftarrow \mathit{EmptyStack}$
\STATE $P \leftarrow \mathit{EmptyStack}$
\ENDFOR
\STATE call SEARCH($\mathit{Entry}$) \\
\STATE ... \\
SEARCH(X) : \\
\FORALL {statements $A$ in $X$} 
\IF {$A$ is a $\phi$-assignment}
   \STATE {remove $A$ and continue}
\ENDIF
\FORALL {variables $V$ in $\mathit{RHS}(A)$}
   \STATE {replace $V$ with $V_i$ where $i = \mathit{Top}(S(V))$}
\ENDFOR

@@

\FORALL {$Y \in Succ(X)$}
\IF {$X$ is a conditional node}
   \STATE {push $\mathit{Cond}, \mathit{Branch}$ onto $P$ // $\mathit{Branch}$ is $T$ if $X \rightarrow Y$ is the True branch, $F$ otherwise}     
\ENDIF   
     \STATE{// $Y$ is not done if $Y$ has an incomplete MUX predecessor or an original CFG predecessor \ana{Better?}}
      \IF {$Y$ has $\phi$-assignments and $Y$ is not done} 
          \STATE {$\mathit{Cond}, \mathit{Branch} \gets$ pop $P$}
          \FORALL {$\phi$-assignments in $Y$}
               \IF {$Z\!: V = \mathit{MUX}(Cond, ..., ...) \in \mathit{Pred}(Y)$}
                  %\STATE {replace $V$ in $RHS(Z)$ with $V_i$, where $i = Top(S(V)$}
                  \STATE {remove edge $X\rightarrow Y$, add edge $X\rightarrow Z$ to $\mathit{CFG}$}
                  \STATE {call SEARCH($Z$) // updates $V$'s in LHS and RHS} 
               \ELSE
                   \STATE {$i = Top(S(V))$}
                   \IF {$\mathit{Branch} == T$}
                      \STATE {create new node $Z\!: V = \mathit{MUX}(\mathit{Cond}, V_i, V)$}
                   \ELSE 
                      \STATE {create new node $Z\!: V = \mathit{MUX}(\mathit{Cond}, V, V_i)$}
                   \ENDIF   
                   \STATE {remove edge $X\rightarrow Y$, add edge $X\rightarrow Z$ to $\mathit{CFG}$} 
               \ENDIF
          \ENDFOR 
      \ENDIF
\ENDFOR
\FORALL {$Y \in \mathit{Children}(X)$}
   \STATE {call SEARCH($Y$)}
\ENDFOR
\FORALL {assignments $A$ in $X$}
   \FORALL {variables $V$ in oldLHS($A$)}
       \STATE {pop S(V)}
   \ENDFOR
\ENDFOR
\end{algorithmic}
\label{alg:algorithm1}
\end{algorithm}

@@

So far we focused on the classical SSA analysis and the extension to create MUX nodes. 
Classical SSA treats arrays as scalars, i.e., an array write, e.g., {\sf A[i] = x+y} is considered
a modification of {\sf A} and warrants its own indexed variable ${\sf A}_k$. A later use of {\sf A}, 
even one such as {\sf x = A[i+1]} where there is no def-use link from {\sf A[i] = x+y}, will be 
considered a use of ${\sf A}_k$. This hinders parallelization. To illustrate, consider our running
example in~\figref{fig:MPCexample}:
\begin{minted}[fontsize=\footnotesize, numbersep=5pt, escapeinside=||]{java}
for (...) {
  A: A[i] = f_1(B[i]);
  B: B[i] = f_2(A[i]);
  ...
}
\end{minted}
Since {\sf A} and {\sf B} are treated as standard scalars, the dependence analysis will record
a dependence from {\sf B} to {\sf A} via the back edge. As a result, statement {\sf A} wouldn't 
be vectorized because each (i+1)-th iteration depends on {\sf B} in the i-th iteration,
which depends on {\sf A} in the i-th iteration.

To mitigate against this, we propose a technique, similar to Array-SSA though much simpler and 
less general, which breaks certain dependence edges and allows for better vectorization. 
\ana{Have to make a claim how 
often this technique resolves dependences perfectly, and how often it defaults to standard SSA.}

@@

\paragraph {Equivalence and Non-equivalence} Let $f(i)$ and $f'(i)$ be two index functions on induction variables $i$, 
e.g., we have array accesses ${\A}[f(i)]$ and ${\A}[f'(i)]$. We say that $f(i) \cong f'(i)$ iff for every $1 \le i \le N$,
$f(i) = f'(i)$. We say that $f(i) \ncong f'(i)$ iff for every $1 \le i \le N$, $f(i) \neq f'(i)$.

\paragraph{Array Well-formedness Conditions} Let ${\A}$ be an array mentioned in the loop body. 
We will treat array accesses ${\A}[f(i)]$ as distinct scalar variables if we can show that the loop body is
well-formed with respect to \A. Otherwise, we treat array {\A} as a single variable, thus defaulting to the 
case of standard SSA. The well-formedness conditions are as follows: \ana{This needs some work!}

@@

\item For each pair $n_1\!: A[f(i)] = ... $ and $n_2\!: ... = A[f'(i)]$ %such that $n_1$ reaches the loop exit (we assume that there is
%a unique loop exit node) on a forward path, and there is  $n_2$ on a forward path, 
we have either $f(i-d) \cong f'(i)$ or $f(i-d) \ncong f'(i)$. \ana{FIX!}

@@

To see the intuition behind these rules, consider rule (1). 
In other words, the two array writes are either equivalent, i.e., at each iteration they modify the same
location, or they are non-equivalent, i.e., at each iteration they modify distinct locations. Note that the two array writes
``clash'' in one way or another. If $n_1$ reaches $n_2$ (or vice versa, $n_2$ reaches $n_1$), then if $f(i)$ and $f'(i)$ 
are neither equivalent nor non-equivalent, there would be a kill in some iterations and not in other; in contrast, if
they are equivalent, there is always a kill, and if they are non-equivalent, there is never a kill. If neither $n_1$ reaches
$n_2$ nor the other way around, then the two array writes will ''clash'' at a join node. If $f(i)$ and $f'(i)$ are neither 
equivalent nor non-equivalent, then there would be one kind of MUX for certain iterations and a different kind of MUX 
for the rest; if $f(i)$ is equivalent to $f'(i)$, then the MUX will be $... = \mathit{MUX}(\mathit{cond}, {\A}[f(i)], {\A[f'(i)]})$, and 
$f(i)$ and $f'(i)$ are non-equivalent then there will be two MUX nodes: $... = \mathit{MUX}(\mathit{cond}, {\A}[f(i)], {\A'[f(i)]})$
and $... = \mathit{MUX}(..., {\A}[f'(i)], {\A'[f'(i)]})$. (Note that after SSA, the A's will be properly indexed.)
\ana{I need explanations of (2) and (3).}

@@

\ana{Where do we put this? Theorem: simple rule is not restrictive, and Theorem 2: program equivalent to transformed one where 
each A[f(i)] is replaced with scalar.}

\subsubsection{Constructing G}

@@

\item We run \ref{alg:algorithm1} to compute SSA of the loop body with MUX nodes. Equivalent accesses get their own indexed versions.

@@

@@

%We will consider Strongly Connected Components (SCCs) in $G$. In the running example in~\figref{fig:MPCexample}(b), 
%we have two SCCs, $SCC_1$ consists of node 2, and $SCC_2$ consists of nodes $3,4,5$. 
Without loss of generality, we may assume that each cycle has the form $n_1 \rightarrow n_2 \rightarrow ... \rightarrow n_k \stackrel{-d}{\dasharrow} n_1$, 
or in other words, there is a single backward edge in the entire cycle. Even though cycles with $n \ge 2$ backward 
edges are possible, such cycles can be transformed into $n$ disjoint cycles of the above form. The $-d$ on the
single backward edge in each of the $n$ cycles, is the sum of the $-d$'s of all backward edges in the original cycle. 
\ana{We probably need more here.}

@@

\item For each SCC, assume that there is a schedule $S: P_1 \rightarrow P_2 ... \rightarrow P_k$ given 
as input. $S$ takes into account only the forward edges in $G$: if $n_1, n_2 \in SCC$ and $n_1 \rightarrow n_2 \in G$, 
then $S(n_1)$ precedes $S(n_2)$ in the linear schedule. Note that computing an optimal schedule is NP-complete, 
as we showed earlier, however, given a small SCC, one can brute force an optimal $S$. For the running example in 
in~\figref{fig:MPCexample}(b), assuming that nodes 3 and 4 perform the same operation, we work with schedule 
$S: \{3,4\} \rightarrow \{5\}$

\item Given $S$ for a given SCC, we then compute the amortized schedule $\mathcal{A}_{\mathit{SCC}}$; 
this is the schedule across all iterations of the loop. For each $P_i$ in $S$, there are $\lceil{\frac{N}{d(SCC)}}\rceil$
nodes in $\mathcal{A}_{\mathit{SCC}}$. $P_i(1),P_i(2),...P_i(d(SCC))$ go into node $P^1_i$, $P_i(d(SCC)+1),...P_i(2\cdot d(SCC))$ 
go into node $P^2_i$, etc. The parenthesized number denotes the iteration, e.g., $P_i(1)$ denotes $P_i$ at the first iteration 
of the loop. There are dependence edges as follows:
\begin{enumerate}
\item $P^j_i \rightarrow P^j_{i'}$ if there is an edge $P_i \rightarrow P_{i'}$ in $S$ (intra-loop edges).
\item $P^{\lceil{\frac{N}{d(SCC)}}\rceil}_i \rightarrow P^1_{i+1}$ for each $1 \le i < \lceil{\frac{N}{d(SCC)}}\rceil$ (across-loop edges).
\end{enumerate}
In the running example, $SCC_1$'s schedule is just the single vectorized node that executes all A's at once.
$SCC_2$'s schedule repeats $S: \{3,4\} \rightarrow \{5\}$ $N$ times: we have $\{3,4\}(1) \rightarrow \{5\}(1) \rightarrow \{3,4\}(2) \rightarrow \{5\}(2) ... \{3,4\}(N) \rightarrow \{5\}(N)$, 
as each iteration depends on the previous-iteration value computed at the same node.
This is illustrated graphically in~\figref{fig:MPCexample}(d).

\item User gives an SCC schedule $SCC_1 \rightarrow SCC_2 \rightarrow ... \rightarrow SCC_m$. Again, an optimal schedule 
is NP-complete. The final schedule is $\mathcal{A} = \mathcal{A}_{\mathit{SCC}_1} \rightarrow \mathcal{A}_{\mathit{SCC}_2} ... \rightarrow \mathcal{A}_{\mathit{SCC}_m}$.
In our running example, there is only one possible ordering of the two components, and the final schedule is as depicted in~\figref{fig:MPCexample}(d).
%dependence edges $P(n)_1 \dasharrow P(n)_2 ... \dasharrow P(n)_{\lceil{\frac{N}{d(n)}}\rceil}$, reflecting the across-loop dependence 
%edges we saw earlier. The rest of the edges for every edge $n_1 \rightarrow n_2 \in G$ and every $1 \le i \le N$, add an edge 
%$G'(n_1(i)) \rightarrow G'(n_2(i))$ to $G'$ where $G'(n(i))$ returns the $P$ node in $G'$ that contains $n(i)$.

@@

\ana{Have to make some kind of claim that certain \% of bench falls under this assumptions or otherwise our results don't count for much.}

\subsection{Theoretical Guarantees}
\label{sec:results}

Unfortunately, finding an optimal across-loop amortized schedule is NP-complete, just as finding an optimal loop-body schedule is.
Our key premise is that one can still compute an optimal across-loop amortized schedule for real-world loops. Given certain 
properties of the loop, the amortized schedule we compute in the previous section turns out to be correct and optimal.

@@

@@

\begin{theorem}(Optimality)
Let $G$ be a constant-rate graph that entails a DAG $D$, where the nodes of $D$ are the strongly connected components of $G$. 
If $D$ is linear, and each $SCC \in D$ entails a loop-body schedule $S: P_1 \rightarrow P_2 ... P_k$, s.t. (1) $S$ is optimal 
and (2) for every $P_i, i<k$, for every $n_1 \in P_i, \exists n_2 \in P_{i+1}$, s.t. $n_1 \rightarrow n_2 \in G$, then $\mathcal{A}$ is optimal.
\end{theorem}

\begin{proof} (Sketch)
Each SCC schedule is optimal because 1) dependences are the same across the loop iterations, and 2) there is dependence between 
iteration $i$ and iteration $i+1$, no node in $i+1$ can be computed before the last node, i.e., $P_k(i)$ has been computed, as all 
nodes in $i+1$ depend on $P_k(i)$. \ana{Expand here.}
\end{proof}

>>>>

\ana{We need an intro here to give a roadmap of what we are going to be doing.}

@@

\begin{figure}
\small
\[
\begin{array}{l@{~~~~~}l}
  \begin{array}{l@{~}l@{~~~}l}
  s & ::= s ; s \\ 
  & \mid \x = \y \; \mid \x=\y \; \op \; \z \mid {\sf A[f(i)] = x} \; \mid {\sf x = A[f(i)]} & \mathit{assignment} \\
%  & \mid \x = \phi(\y,\z) \\
  & \mid {\sf for} \; (i = 0; i \le \mathit{Const}; \mathit{i\!+\!+}) \; \{ \; s \; \} \\
  & \mid {\sf if} \; (\x \; \cop \; \y) \; {\sf then} \; \{ \; s \; \} \; {\sf else} \; \{ \; s \; \} \\
 % \bool &::= \x \; \cop \; \y & \mathit{comparison} \\
  \op &::= + \mid - \mid * \mid / & \mathit{arithmetic} \; \mathit{operator} \\
  \cop &::=  \; == \; \mid \; != \; \mid \; < \; \mid \; \le & \mathit{comparison} \; \mathit{operator}

@@

\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
z = 0;
...
if (x > 0) {
  if (y > 0) {
    z = 1;
  }  else {
    z = 2;
  }
}
print z
\end{minted}
Line 10 is the dominance frontier of the assignments to \z, and Cytron et. al.'s algorithm
creates a $\phi$ node at 10: $z_3 = \phi(z_0, z_1, z_2)$. There is no information retained
that $z_0$ is triggered under path condition $x \le 0$, $z_1$ is triggered under path condition 
$x > 0 \; \wedge \; y > 0$, and $z_2$ is triggered when $x > 0 \; \wedge \; y \le 0$. The 
Multiplexer at the join point needs this information however in order to assign the correct 
value to $z_3$ at runtime.

@@

\begin{algorithm}
\caption{From $\phi$-nodes to MUX nodes}
\begin{algorithmic}[1]
%\REQUIRE $c \neq null$
%\STATE $L \leftarrow get\_nodes(c)$
%\FORALL{$n$ in $L$}
%\STATE $...$
%\ENDFOR
\FORALL {variables $V$}
\STATE $C(V) \leftarrow 0$
\STATE $S(V) \leftarrow \mathit{EmptyStack}$
\STATE $P \leftarrow \mathit{EmptyStack}$
\ENDFOR
\STATE call SEARCH($\mathit{Entry}$) \\
\STATE ... \\
SEARCH(X) : \\
\FORALL {statements $A$ in $X$} 
\IF {$A$ is a $\phi$-assignment}
   \STATE {remove $A$ and continue}
\ENDIF
\FORALL {variables $V$ in $\mathit{RHS}(A)$}
   \STATE {replace $V$ with $V_i$ where $i = \mathit{Top}(S(V))$}
\ENDFOR

@@

\FORALL {$Y \in Succ(X)$}
\IF {$X$ is a conditional node}
   \STATE {push $\mathit{Cond}, \mathit{Branch}$ onto $P$ // $\mathit{Branch}$ is $T$ if $X \rightarrow Y$ is the True branch, $F$ otherwise}     
\ENDIF   
     \STATE{// $Y$ is not done if $Y$ has an incomplete MUX predecessor or an original CFG predecessor}
      \IF {$Y$ has $\phi$-assignments and $Y$ is not done} 
          \STATE {$\mathit{Cond}, \mathit{Branch} \gets$ pop $P$}
          \FORALL {$\phi$-assignments in $Y$}
               \IF {$Z\!: V = \mathit{MUX}(Cond, ..., ...) \in \mathit{Pred}(Y)$}
                  %\STATE {replace $V$ in $RHS(Z)$ with $V_i$, where $i = Top(S(V)$}
                  \STATE {remove edge $X\rightarrow Y$, add edge $X\rightarrow Z$ to $\mathit{CFG}$}
                  \STATE {call SEARCH($Z$) // updates $V$'s in LHS and RHS} 
               \ELSE
                   \STATE {$i = Top(S(V))$}
                   \IF {$\mathit{Branch} == T$}
                      \STATE {create new node $Z\!: V = \mathit{MUX}(\mathit{Cond}, V_i, V)$}
                   \ELSE 
                      \STATE {create new node $Z\!: V = \mathit{MUX}(\mathit{Cond}, V, V_i)$}
                   \ENDIF   
                   \STATE {remove edge $X\rightarrow Y$, add edge $X\rightarrow Z$ to $\mathit{CFG}$} 
               \ENDIF
          \ENDFOR 
      \ENDIF
\ENDFOR
\FORALL {$Y \in \mathit{Children}(X)$}
   \STATE {call SEARCH($Y$)}
\ENDFOR
\FORALL {assignments $A$ in $X$}
   \FORALL {variables $V$ in oldLHS($A$)}
       \STATE {pop S(V)}
   \ENDFOR
\ENDFOR
\end{algorithmic}
\label{alg:algorithm1}
\end{algorithm}

@@

So far we focused on the classical SSA analysis and the extension necessary to create MUX nodes. 
Classical SSA treats arrays as scalars, i.e., an array write, e.g., {\sf A[i] = x+y} is considered
a modification of {\sf A} and warrants its own indexed variable ${\sf A}_k$. A subsequent use of {\sf A}, 
even one such as {\sf x = A[i+1]} where there is no def-use link from {\sf A[i] = x+y}, will be 
considered a use of ${\sf A}_k$. This hinders parallelization. To illustrate, consider our running
example in~\figref{fig:MPCexample}:
\begin{minted}[fontsize=\footnotesize, numbersep=5pt, escapeinside=||]{java}
for (...) {
  A: A[i] = f_1(B[i]);
  B: B[i] = f_2(A[i]);
  ...
}
\end{minted}
Since {\sf A} and {\sf B} are treated as standard scalars, the dependence analysis will record
a dependence from {\sf B} to {\sf A} via the back edge. As a result, statement {\sf A} wouldn't 
be vectorized because each (i+1)-th iteration depends on {\sf B} in the i-th iteration,
which depends on {\sf A} in the i-th iteration.

To mitigate against this, we propose a technique, similar to Array-SSA though much simpler, 
which breaks certain dependence edges and allows for better vectorization. 
\ana{Have to make a claim how 
often this technique resolves dependences perfectly, and how often it defaults to standard SSA.}

@@

\paragraph {Equivalence and Non-equivalence} Let $f(i)$ and $f'(i)$ be two index functions on induction variable $i$, 
e.g., we have array accesses ${\A}[f(i)]$ and ${\A}[f'(i)]$. We say that $f(i) \cong f'(i)$ iff for every $1 \le i \le N$,
$f(i) = f'(i)$. We say that $f(i) \ncong f'(i)$ iff for every $1 \le i \le N$, $f(i) \neq f'(i)$.

\ana{This needs explanation and examples. The context is single loops where multidimensional arrays are turned 
into one dimensional ones. The index function becomes complex.}

\paragraph{Array Well-formedness Conditions} Let ${\A}$ be an array mentioned in the loop body. 
We treat array accesses ${\A}[f(i)]$ as distinct scalar variables if we can show that the loop body is
well-formed with respect to \A. Otherwise, we treat array {\A} as a single variable, thus defaulting to the 
case of standard SSA. The well-formedness conditions are as follows: \ana{This needs some work!}

@@

\item For each pair $n_1\!: A[f(i)] = ... $ and $n_2\!: ... = A[f'(i)]$ %such that $n_1$ reaches the loop exit (we assume that there is
%a unique loop exit node) on a forward path, and there is  $n_2$ on a forward path, 
we have $f(i-d) \cong f'(i)$ for at most one $d$ such that $1 \le d \le N-1$ and 
we have $f(i-d') \ncong f'(i)$ for every $d' \neq d$.

@@

To see the intuition behind these rules, consider rule (1). 
In other words, the two array writes are either equivalent, i.e., at each iteration they modify the same
location, or they are non-equivalent, i.e., at each iteration they modify distinct locations. Note that the two array writes
``clash'' in one way or another. If $n_1$ reaches $n_2$ (or vice versa, $n_2$ reaches $n_1$), then if $f(i)$ and $f'(i)$ 
are neither equivalent nor non-equivalent, there would be a kill in some iterations and not in other; in contrast, if
they are equivalent, there is always a kill, and if they are non-equivalent, there is never a kill. If neither $n_1$ reaches
$n_2$ nor the other way around, then the two array writes will ''clash'' at a join node. If $f(i)$ and $f'(i)$ are neither 
equivalent nor non-equivalent, then there would be one kind of MUX for certain iterations and a different kind of MUX 
for the rest; if $f(i)$ is equivalent to $f'(i)$, then the MUX will be $... = \mathit{MUX}(\mathit{cond}, {\A}[f(i)], {\A[f'(i)]})$, and 
$f(i)$ and $f'(i)$ are non-equivalent then there will be two MUX nodes: $... = \mathit{MUX}(\mathit{cond}, {\A}[f(i)], {\A'[f(i)]})$
and $... = \mathit{MUX}(..., {\A}[f'(i)], {\A'[f'(i)]})$. (Note that after SSA, the A's will be properly indexed.)
\ana{I need explanations of (2) and (3).}
Now consider rule (3). If there exist a $d$ such that $f(i-d) \cong f'(i)$, that means that the array write $j: \A[f(i)] = ... $ that
happened $d$ iterations before array read $k: ... = \A[f'(i)]$ is being used at that read. In addition, this holds for every $i$, 
i.e., for every iteration, the array read at $k$ references the array write at $j$ that happened $d$ iterations ago. Also,
since for every other $d'$ the array accesses are non-equivalent, this means, that the read at $k$ never references  
the write at $j$ in a different iteration.
%any other previous iteration, just $I-d$, where $I$ is the current iteration.

@@

\subsubsection{Constructing MPC G}

@@

\item We run Algorithm~\ref{alg:algorithm1} to compute SSA of the loop body with MUX nodes. Equivalent accesses get their own indexed versions.

@@

\ana{Here we need a theorem: program equivalent to transformed one in which every A[f(i)] is replaced with scalar. 
Prove that original def-use graph and one with scalars are the same?}

@@

%We will consider Strongly Connected Components (SCCs) in $G$. In the running example in~\figref{fig:MPCexample}(b), 
%we have two SCCs, $SCC_1$ consists of node 2, and $SCC_2$ consists of nodes $3,4,5$. 
Without loss of generality, we may assume that each cycle has the form $n_1 \rightarrow n_2 \rightarrow ... \rightarrow n_k \stackrel{-d}{\dasharrow} n_1$, 
or in other words, there is a single backward edge in the entire cycle. Even though cycles with more than $2$ backward 
edges are possible, such cycles can be transformed into $n$ disjoint cycles of the above form. The $-d$ on the
single backward edge in each of the $n$ cycles, is the sum of the $-d$'s of all backward edges in the original cycle. 
\ana{We probably need more here.}

@@

\item For each SCC, assume that there is a schedule $S: P_1 \rightarrow P_2 ... \rightarrow P_k$ given 
as input. $S$ takes into account only the forward edges in $G$: if $n_1, n_2 \in SCC$ and $n_1 \rightarrow n_2 \in G$, 
then $S(n_1)$ precedes $S(n_2)$ in the linear schedule. Note that computing an optimal schedule is NP-complete, 
as we showed earlier, however, given a small SCC, one can brute force an optimal $S$. For the running example in 
~\figref{fig:MPCexample}(b), assuming that nodes 3 and 4 perform the same operation, we work with schedule 
$S: \{3,4\} \rightarrow \{5\}$. Importantly, due to our treatment of arrays we have the same schedule for 
each iteration of the loop, i.e., for each iteration, we either have a dependence edge or we do not have a dependence.

\item Given $S$ for a given SCC, we then compute the amortized schedule $\mathcal{A}_{\mathit{SCC}}$; 
this is the schedule across all iterations of the loop. For each $P_i$ in $S$, there are $\lceil{\frac{N}{d(SCC)}}\rceil$
nodes in $\mathcal{A}_{\mathit{SCC}}$:
\begin{itemize}
\item[] $P_i(1),P_i(2),...P_i(d(SCC))$ go into node $P^1_i$, 
\item[] $P_i(d(SCC)+1),...P_i(2\cdot d(SCC))$ go into node $P^2_i$, etc. 
\end{itemize}
The parenthesized number denotes the iteration, e.g., $P_i(1)$ denotes $P_i$ at the first iteration 
of the loop.

There are dependence edges as follows:
\begin{enumerate}
\item $P^j_i \rightarrow P^j_{i'}$ if there is an edge $P_i \rightarrow P_{i'}$ in $S$ (intra-loop edges).
\item $P^{\lceil{\frac{N}{d(SCC)}}\rceil}_i \rightarrow P^1_{i+1}$ for each $1 \le i < \lceil{\frac{N}{d(SCC)}}\rceil$ (across-loop edges).
\end{enumerate}
In the running example, $SCC_1$'s schedule is just the single vectorized node that executes all A's at once.
$SCC_2$'s schedule repeats $S: \{3,4\} \rightarrow \{5\}$ $N$ times: we have $\{3,4\}(1) \rightarrow \{5\}(1) \rightarrow \{3,4\}(2) \rightarrow \{5\}(2) ... \{3,4\}(N) \rightarrow \{5\}(N)$, 
as each iteration depends on the previous-iteration value computed at the same node.
This is illustrated graphically in~\figref{fig:MPCexample}(d).

\item Computing an optimal schedule for the loop, even when we assume optimal schedules 
for the SCCs, remains NP-complete. 
One possible schedule can be computed by ordering the strongly connected components 
in a straight line: $SCC_1 \rightarrow SCC_2 \rightarrow ... \rightarrow SCC_m$, such that
if $SCC_i$ precedes $SCC_j$ in the SCC DAG, then $SCC_i$ precedes $SCC_j$ in the
above straight line. The final schedule is $\mathcal{A} = \mathcal{A}_{\mathit{SCC}_1} 
\rightarrow \mathcal{A}_{\mathit{SCC}_2} ... \rightarrow \mathcal{A}_{\mathit{SCC}_m}$.
In our running example, there is only one possible ordering of the two components, and 
the final schedule is as depicted in~\figref{fig:MPCexample}(d).

@@

\subsection{Theoretical Guarantees on Correctness and Optimality}
\label{sec:results}

Unfortunately, finding an optimal across-loop amortized schedule is NP-complete, just as finding an optimal loop-body schedule is.
Our key premise is that the schedule we compute in~\secref{sec:scc} (1) improves upon the original schedule $\mathcal{S}$, and 
(2) is an optimal across-loop amortized schedule in many cases of real-world loop examples. \ana{Talk about decomposition.}

@@

\ana{1. Theorem: Each SCC schedule is optimal}

\ana{2. Theorem: An optimal across-loop schedule (even under assumptions about optimality of individual SCC schedules) is NP-complete.}

\ana{Conditions under which schedule is optimal. Exploit the fact that all SCCs are either d=1 or d=N. Scheduling SCCs in parallel and computing the schedule.
Also, ???}

@@

%\begin{theorem}(Optimality)
%Let $G$ be a constant-rate graph that entails a DAG $D$, where the nodes of $D$ are the strongly connected components of $G$. 
%If $D$ is linear, and each $SCC \in D$ entails a loop-body schedule $S: P_1 \rightarrow P_2 ... P_k$, s.t. (1) $S$ is optimal 
%and (2) for every $P_i, i<k$, for every $n_1 \in P_i, \exists n_2 \in P_{i+1}$, s.t. $n_1 \rightarrow n_2 \in G$, then $\mathcal{A}$ is optimal.
%\end{theorem}

%\begin{proof} (Sketch)
%Each SCC schedule is optimal because 1) dependences are the same across the loop iterations, and 2) there is dependence between 
%iteration $i$ and iteration $i+1$, no node in $i+1$ can be computed before the last node, i.e., $P_k(i)$ has been computed, as all 
%nodes in $i+1$ depend on $P_k(i)$. \ana{Expand here.}
%\end{proof}
